---
title: "OGP 2020 Additional Scenarios"
author: "Neal Fultz <<neal@njnm.co>>"
date: "4/22/2020"
output: html_vignette
vignette: >
  %\VignetteIndexEntry{OGP 2020 Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(ogp)

knitr::opts_chunk$set(echo=FALSE)

if(! "Current.Awards" %in% colnames(allocations)) {
  allocations <- cbind(
    allocations,
    Current = Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year)
  )
}

```






## Scenarios

```{r scenarios, error=FALSE, results='hide'}

adjustFinalCategoryToConstraint <- function(scen, limit) {
  loss <- function(x) abs(limit - OGP.2013_forward(40000000, 
        scen - c(0,0,0, 0,0,0, x,x,0)))
  x <- optim(0, loss, method="Brent", lower=-.01, upper=scen[8])
  scen - c(0,0,0,0,0,0,x$par,x$par,0)
}


# Scenario 5 - 10% flat reduction
scenario5 <- scenario(allocations,
                      Y_low = .90 * allocations$Current.Grant100,
); .Last.value
#scenario5 <- c(6750, 0, 0, 0.3545, 0.3545, 0.0337, 0.0049, 0.0049, 0)

scenario5 <- adjustFinalCategoryToConstraint(scenario5, 270000)


ineq1_cap <- matrix(ncol=10, byrow=TRUE, -c(
1, 7499, 10000, 32500, 50000, 1400000, 8500000, 3e+07, 0, 250000
)/100000)

ineq2_cap <- matrix(ncol=10, byrow=TRUE, -c(
1, 7499, 10000, 32500, 50000, 1400000, 8500000, 3e+07, 0, 200000
)/100000)


ineq1_base <- matrix(ncol=10, byrow=TRUE, c(
1, 0, 0, 0, 0, 0, 0, 0, 0,    6750
))


eq5a <- matrix(ncol=10, byrow=TRUE, c(
0, 1, 0, 0, 0, 0, 0, 0, 0,    0,
0, 0, 1, 0, 0, 0, 0, 0, 0,    0,
0, 0, 0,-1, 1, 0, 0, 0, 0,    0,
0, 0, 0, 0, 0, 0,-1, 1, 0,    0,
0, 0, 0, 0, 0, 0, 0, 0, 1,    0
))



scenario5a <- scenario(allocations,
                      Y_low = .90 * allocations$Current.Grant100,
                      mod=list(extra_eq=eq5a, extra_inq=rbind(ineq1_cap,ineq1_base),
                               method='LU'),
                      groups_const = c("City", "District_Most_Activity", "Discipline"),
                      niter=100#, verbose=TRUE
); .Last.value
scenario5a[1] <- 6750

#scenario5a <- c(6750, 0, 0, 0.3961, 0.3961, 0.0385, 0.0041, 0.0041, 0)
scenario5a <- adjustFinalCategoryToConstraint(scenario5a, 250000)


scenario5b <- scenario(allocations,
                      Y_low = .90 * allocations$Current.Grant100,
                      mod=list(extra_eq=eq5a, extra_inq=rbind(ineq2_cap,ineq1_base),
                               s=1.2, 
                               base=90000,
                               method='PFCF'),
                      groups_const = c("City", "District_Most_Activity", "Discipline"),
                      niter=86); .Last.value

scenario5b[1] <- 6750
scenario5b[4:5] <- mean(scenario5b[4:5])
scenario5b[7:8] <- mean(scenario5b[7:8])

#scenario5b <- c(6750, 0, 0, 0.3852, 0.3852, 0.056, 0.0029, 0.0029, 0)
scenario5b <- adjustFinalCategoryToConstraint(scenario5b, 200000)


#print(list(s, scenario5b))
#}

# scenario5b <- c(7500, 0, 0, 0.3862, 0.3862, 0.0386, 0.0041, 0.0041, 0)


### Scenario 6 - 100% match for micro up to $7500, recalibrate
eq23 <- matrix(ncol=10, byrow=TRUE, c(
0, 0, -1,  1, 0, 0, 0, 0, 0,   0,
0, 0, -1,  0, 1, 0, 0, 0, 0,   0,
0, 0,  0,  0, 0, 0, 1,-1, 0,   0,  
0, 0,  0,  0, 0, 0, 0, 0,-1,   0
))


ineq23 <- matrix(ncol=10, byrow=TRUE, c(
0, 1, 0,  0, 0, 0, 0, 0, 0,   .99
))



scenario6 <- scenario(allocations,
                      Y_low = .90 * allocations$Current.Grant100,
                      mod=list(base=3, extra_eq=eq23, extra_inq=ineq23,
                               s=0.48, method='PFCF'),
                      niter=43); .Last.value
scenario6[1:2] <- 0:1
#scenario6 <- c(0, 1, 0.3279, 0.3279, 0.3279, 0.0325, 0.0049, 0.0049, 0)
  
scenario6a <- scenario(allocations,
                      Y_low = .90 * allocations$Current.Grant100,
                      groups_const = c("City", "District_Most_Activity", "Discipline"),
                      mod=list(base=3, extra_eq=eq23, extra_inq=rbind(ineq23,ineq1_cap),
                               s=0.49, method='LU'),
                      niter=82#, verbose=TRUE
                      ); .Last.value
scenario6a[1:2] <- 0:1
#scenario6a <- c(0, 1, 0.3312, 0.3312, 0.3312, 0.0407, 0.004, 0.004, 0)
scenario6a <- adjustFinalCategoryToConstraint(scenario6a, 250000)


# #for(niter in seq(130,200, 5)) {
# niter=80
# for(s1 in seq(.8, 10, .1)){
# for(s2 in seq(1.3, 10, .1)){
# for(s in seq(.8911, 1.4813, .01)){
# 
# #s <- 0.48338; niter=80  
#   capture.output(
#   scenario6b <- scenario(allocations,
#                       Y_low = .90 * allocations$Current.Grant100,
#                       groups_const = c("City", "District_Most_Activity", "Discipline"),
#                       mod=list(extra_eq=eq23, extra_inq=rbind(ineq23/s1,ineq2_cap/s2),
#                                s=s,#s=.4805, 
#                                base=100,
#                                method='PFCF'),
#                       niter=niter)#; .Last.value
#   )
# 
#   if(!is.finite(scenario6b[1])) next
#   if(scenario6b[2] > .99) {
#     print("------------------")
#     print(list(s=c(s,s1,s2), niter=niter, scenario6b))
#   
#   }
#   if(scenario6b[2] > .999 && 
#      diff(range(scenario6b[3:5])) < .005 &&  
#      diff(range(scenario6b[7:8])) < .0005
#      ) {stop("found")}
# }}}


s=1.2211; s1=0.8000; s2=1.8000; niter=80;
  scenario6b <- scenario(allocations,
                      Y_low = .90 * allocations$Current.Grant100,
                      groups_const = c("City", "District_Most_Activity", "Discipline"),
                      mod=list(extra_eq=eq23, extra_inq=rbind(ineq23/s1,ineq2_cap/s2),
                               s=s,#s=.4805, 
                               base=100,
                               method='PFCF'),
                      niter=niter+7 ); .Last.value
scenario6b[1:2] <- 0:1
scenario6b[3:5] <- mean(scenario6b[3:5])
scenario6b[7:8] <- mean(scenario6b[7:8])
#scenario6b <- c(0, 1, 0.3958, 0.3958, 0.3958, 0.0325, 0.003, 0.003, 0)

scenario6b <- adjustFinalCategoryToConstraint(scenario6b, 200000)




```

By applying extra constraints around the $1 award, the micro category slope, and the category 1 slope, we can create a set of
several different solutions:
  
1. Scenario 5
    - No additional constraints
    - In this scenario, all Max Requests are reduced by 10%.
    - Because the awards are ultimately rescaled to the County Funds, little impact on final awards or Percent Awarded.
2. Scenario 5a 
    - as Scenario 5, but max request is $250,000
3. Scenario 5b 
    - as Scenario 5, but max request is $200,000
4. Scenario 6 
    - as Scenario 5, but 100% match up to $7500 instead of flat
2. Scenario 6a 
    - as Scenario 5a, but 100% match up to $7500 instead of flat
3. Scenario 6b 
    - as Scenario b, but 100% match up to $7500 instead of flat

We can evaluate these scenarios at various budget sizes. All scenarios look very similar for the high end, but can be quite different at the small end:

```{r ScenComp}
b <- cumsum(basis(100*1000000))
kable(cbind(
  Budget = b, 
  Current= OGP.2013_forward(b),
  Scenario5= OGP.2013_forward(b, scenario5),
  Scenario5a= OGP.2013_forward(b, scenario5a),
  Scenario5b= OGP.2013_forward(b, scenario5b),
  Scenario6= OGP.2013_forward(b, scenario6),
  Scenario6a= OGP.2013_forward(b, scenario6a),
  Scenario6b= OGP.2013_forward(b, scenario6b)
), digits = -2, format.args=list(big.mark=',', scientific = FALSE), caption="Max Request Amounts") %>% column_spec(1, border_right = TRUE)

```



```{r error=FALSE, warning=FALSE, fig.height=6, fig.width=8, fig.align="center", out.width="80%"}
zzz <- data.frame(Budget=b,
  Current= OGP.2013_forward(b),
  Scenario5= OGP.2013_forward(b, scenario5),
  Scenario5a= OGP.2013_forward(b, scenario5a),
  Scenario5b= OGP.2013_forward(b, scenario5b),
  Scenario6= OGP.2013_forward(b, scenario6),
  Scenario6a= OGP.2013_forward(b, scenario6a),
  Scenario6b= OGP.2013_forward(b, scenario6b)
)

zzz_long <- melt(zzz, id.vars="Budget", variable.name="Scenario", value.name = "MaxAward")

ggplot(zzz_long) + aes(x=Budget, y=MaxAward, color=Scenario) + geom_line() +
  geom_point(aes(x=Budget_Size), data=allocations, y=0, pch='x', color='black') + 
  scale_y_continuous(name = "Max Request", limits = c(0,300000), labels = scales::dollar) + 
  scale_x_continuous(name = "Budget", limits = c(1,100000000), labels = scales::dollar) +
  ggtitle("Max Request by Budget")

```

```{r error=FALSE, warning=FALSE, fig.height=6, fig.width=8, fig.align="center", out.width="80%"}
# ggplot(zzz_long) + aes(x=Budget, y=MaxAward, color=Scenario) + geom_line(aes(linetype=Scenario != "Current"), size=1.5) + xlim(1, 50000) + ylim(0, 30000) +
#   geom_point(aes(x=Budget_Size), data=allocations, y=0, pch='x', color='black') + guides(linetype="none") + ggtitle("OGP I Scenarios")


ggplot(zzz_long) + aes(x=Budget, y=MaxAward, color=Scenario) + geom_line() +
  geom_point(aes(x=Budget_Size), data=allocations, y=0, pch='x', color='black') + 
  scale_y_continuous(name="Max Request", limits = c(0,30000), labels = scales::dollar) + 
  scale_x_continuous(name="Budget", limits = c(1,50000), labels = scales::dollar) +
  ggtitle("Max Request by Budget (Smaller categories)") + 
  geom_abline(slope=1, intercept=0, linetype='dashed', alpha=.4)

```

# Comparison by final awards (4.5M Total County Budget) by group
```{r all_sec_counterfactual}
  Current  = Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=NULL)
  Scenario5= Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=scenario5)
  Scenario5a= Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=scenario5a)
  Scenario5b= Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=scenario5b)
  Scenario6= Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=scenario6)
  Scenario6a= Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=scenario6a)
  Scenario6b= Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=scenario6b)

  all_scenarios <- cbind( allocations[c('Organization', 'Year', 'OGP_Budget_Category', "District_Most_Activity", "Discipline", "Budget_Size")],
                          Current=Current, 
                          Scenario5=Scenario5, 
                          Scenario5a=Scenario5a, 
                          Scenario5b=Scenario5b,
                          Scenario6=Scenario6, 
                          Scenario6a=Scenario6a, 
                          Scenario6b=Scenario6b)
  
```

### Change in funding rate


```{r}
kable(rbind.data.frame(
  data.frame(Scenario="Current", `Total Scored Request` = sum(Current$Grant100), `Total Final Award`=sum(Current$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`),
data.frame(Scenario="Scenario5", `Total Scored Request` = sum(Scenario5$Grant100), `Total Final Award`=sum(Scenario5$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`),
data.frame(Scenario="Scenario5a", `Total Scored Request` = sum(Scenario5a$Grant100), `Total Final Award`=sum(Scenario5a$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`),
data.frame(Scenario="Scenario5b", `Total Scored Request` = sum(Scenario5b$Grant100), `Total Final Award`=sum(Scenario5b$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`),
data.frame(Scenario="Scenario6", `Total Scored Request` = sum(Scenario6$Grant100), `Total Final Award`=sum(Scenario6$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`),
data.frame(Scenario="Scenario6a", `Total Scored Request` = sum(Scenario6a$Grant100), `Total Final Award`=sum(Scenario6a$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`),
data.frame(Scenario="Scenario6b", `Total Scored Request` = sum(Scenario6b$Grant100), `Total Final Award`=sum(Scenario6b$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`)
) %>%   mutate(p=round(p,3), `Total Final Award` = NULL, `Total Scored Request` = round(`Total Scored Request`, -2)), 
               format.args=list(big.mark=',')) 
```


#### Largest changes in each scenario

  
```{r}
all_scenarios %>% mutate(Scenario=Scenario5.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
  arrange(-Budget_Size) %>% filter(rank(-abs(percentChange))< 20 | rank(-abs(delta))< 20 ) %>% 
 select(Organization, Budget_Size, Current.Final, Scenario, delta, percentChange) %>% 
  kable(digits = 2, format.args=list(big.mark=','), caption="Largest Changes in Final Awards, Scenario5")

```
  
```{r}
all_scenarios %>% mutate(Scenario=Scenario5a.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
  arrange(-Budget_Size) %>% filter(rank(-abs(percentChange))< 20 | rank(-abs(delta))< 20 ) %>% 
 select(Organization, Budget_Size, Current.Final, Scenario, delta, percentChange) %>% 
  kable(digits = 2, format.args=list(big.mark=','), caption="Largest Changes in Final Awards, Scenario5a")

```

```{r}
all_scenarios %>% mutate(Scenario=Scenario5b.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
  arrange(-Budget_Size) %>% filter(rank(-abs(percentChange))< 20 | rank(-abs(delta))< 20 ) %>% 
 select(Organization, Budget_Size, Current.Final, Scenario, delta, percentChange) %>% 
  kable(digits = 2, format.args=list(big.mark=','), caption="Largest Changes in Final Awards, Scenario5b")

```


```{r}
all_scenarios %>% mutate(Scenario=Scenario6.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
  arrange(-Budget_Size) %>% filter(rank(-abs(percentChange))< 20 | rank(-abs(delta))< 20 ) %>% 
 select(Organization, Budget_Size, Current.Final, Scenario, delta, percentChange) %>% 
  kable(digits = 2, format.args=list(big.mark=','), caption="Largest Changes, Scenario6")

```

```{r}
all_scenarios %>% mutate(Scenario=Scenario6a.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
  arrange(-Budget_Size) %>% filter(rank(-abs(percentChange))< 20 | rank(-abs(delta))< 20 ) %>% 
 select(Organization, Budget_Size, Current.Final, Scenario, delta, percentChange) %>% 
  kable(digits = 2, format.args=list(big.mark=','), caption="Largest Changes, Scenario6a")

```

```{r}
all_scenarios %>% mutate(Scenario=Scenario6b.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
  arrange(-Budget_Size) %>% filter(rank(-abs(percentChange))< 20 | rank(-abs(delta))< 20 ) %>% 
 select(Organization, Budget_Size, Current.Final, Scenario, delta, percentChange) %>% 
  kable(digits = 2, format.args=list(big.mark=','), caption="Largest Changes, Scenario6b")

```


```{r}
#all_scenarios %>% mutate(Scenario=Scenario2.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
#  ggplot() + aes(x=percentChange) + geom_histogram()
```


#### Total changes

We can also aggregate the changes in final reward by whether it increased or not, and by how much:


```{r deltatbl}
zzz <- all_scenarios[c('Scenario5.Final','Scenario5a.Final','Scenario5b.Final','Scenario6.Final', 'Scenario6a.Final','Scenario6b.Final')] - all_scenarios$Current.Final

zzz <- data.frame(Organization=all_scenarios$Organization, OGP_Budget_Category=all_scenarios$OGP_Budget_Category, zzz)


melt(zzz, c("Organization", "OGP_Budget_Category")) %>% 
  group_by(Scenario=sub(".Final", "", variable), 
           delta=ifelse(sign(value) > 0, '\\+', '\\-')
           ) %>% 
  summarise(n=n(), avg=mean(value), min=min(value), med=median(value), max=max(value)) %>% 
    kable( digits=0, caption="Types / Sizes of Changes")

```

```{r deltatbl2}

melt(zzz, c("Organization", "OGP_Budget_Category")) %>% 
  group_by(Scenario=sub(".Final", "", variable), 
           `Formula Group`=OGP_Budget_Category,
           delta=ifelse(sign(value) > 0, '\\+', '\\-')
           ) %>% 
  summarise(n=n(), avg=mean(value), min=min(value), med=median(value), max=max(value)) %>% 
    kable( digits=0, caption="Types / Sizes of Changes by Formula Group",align='lccrrrrr') %>% 
    row_spec(c(7:13, 20:25, 33:38), background="lightgrey")

```
  


# Comparison by final awards (3.1M Total County Budget) by group
```{r all_sec_counterfactual31}
  Current  = Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=NULL, TotalBudget = 3100000)
  Scenario5= Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=scenario5, TotalBudget = 3100000)
  Scenario5a= Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=scenario5a, TotalBudget = 3100000)
  Scenario5b= Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=scenario5b, TotalBudget = 3100000)
  Scenario6= Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=scenario6, TotalBudget = 3100000)
  Scenario6a= Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=scenario6a, TotalBudget = 3100000)
  Scenario6b= Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=scenario6b, TotalBudget = 3100000)

  all_scenarios <- cbind( allocations[c('Organization', 'Year', 'OGP_Budget_Category', "District_Most_Activity", "Discipline", "Budget_Size")],
                          Current=Current, 
                          Scenario5=Scenario5, 
                          Scenario5a=Scenario5a, 
                          Scenario5b=Scenario5b,
                          Scenario6=Scenario6, 
                          Scenario6a=Scenario6a, 
                          Scenario6b=Scenario6b)
  
```

### Change in funding rate


```{r}
kable(rbind.data.frame(
  data.frame(Scenario="Current", `Total Scored Request` = sum(Current$Grant100), `Total Final Award`=sum(Current$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`),
data.frame(Scenario="Scenario5", `Total Scored Request` = sum(Scenario5$Grant100), `Total Final Award`=sum(Scenario5$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`),
data.frame(Scenario="Scenario5a", `Total Scored Request` = sum(Scenario5a$Grant100), `Total Final Award`=sum(Scenario5a$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`),
data.frame(Scenario="Scenario5b", `Total Scored Request` = sum(Scenario5b$Grant100), `Total Final Award`=sum(Scenario5b$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`),
data.frame(Scenario="Scenario6", `Total Scored Request` = sum(Scenario6$Grant100), `Total Final Award`=sum(Scenario6$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`),
data.frame(Scenario="Scenario6a", `Total Scored Request` = sum(Scenario6a$Grant100), `Total Final Award`=sum(Scenario6a$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`),
data.frame(Scenario="Scenario6b", `Total Scored Request` = sum(Scenario6b$Grant100), `Total Final Award`=sum(Scenario6b$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`)
) %>%   mutate(p=round(p,3), `Total Final Award` = NULL, `Total Scored Request` = round(`Total Scored Request`, -2)), 
               format.args=list(big.mark=',')) 
```


#### Largest changes in each scenario

Note the Current Column displayed below is the Current OGP Formula with a 3.1M Budget.

```{r}
all_scenarios %>% mutate(Scenario=Scenario5.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
  arrange(-Budget_Size) %>% filter(rank(-abs(percentChange))< 20 | rank(-abs(delta))< 20 ) %>% 
 select(Organization, Budget_Size, Current.Final, Scenario, delta, percentChange) %>% 
  kable(digits = 2, format.args=list(big.mark=','), caption="Largest Changes in Final Awards, Scenario5")

```
  
```{r}
all_scenarios %>% mutate(Scenario=Scenario5a.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
  arrange(-Budget_Size) %>% filter(rank(-abs(percentChange))< 20 | rank(-abs(delta))< 20 ) %>% 
 select(Organization, Budget_Size, Current.Final, Scenario, delta, percentChange) %>% 
  kable(digits = 2, format.args=list(big.mark=','), caption="Largest Changes in Final Awards, Scenario5a")

```

```{r}
all_scenarios %>% mutate(Scenario=Scenario5b.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
  arrange(-Budget_Size) %>% filter(rank(-abs(percentChange))< 20 | rank(-abs(delta))< 20 ) %>% 
 select(Organization, Budget_Size, Current.Final, Scenario, delta, percentChange) %>% 
  kable(digits = 2, format.args=list(big.mark=','), caption="Largest Changes in Final Awards, Scenario5b")

```


```{r}
all_scenarios %>% mutate(Scenario=Scenario6.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
  arrange(-Budget_Size) %>% filter(rank(-abs(percentChange))< 20 | rank(-abs(delta))< 20 ) %>% 
 select(Organization, Budget_Size, Current.Final, Scenario, delta, percentChange) %>% 
  kable(digits = 2, format.args=list(big.mark=','), caption="Largest Changes, Scenario6")

```

```{r}
all_scenarios %>% mutate(Scenario=Scenario6a.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
  arrange(-Budget_Size) %>% filter(rank(-abs(percentChange))< 20 | rank(-abs(delta))< 20 ) %>% 
 select(Organization, Budget_Size, Current.Final, Scenario, delta, percentChange) %>% 
  kable(digits = 2, format.args=list(big.mark=','), caption="Largest Changes, Scenario6a")

```

```{r}
all_scenarios %>% mutate(Scenario=Scenario6b.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
  arrange(-Budget_Size) %>% filter(rank(-abs(percentChange))< 20 | rank(-abs(delta))< 20 ) %>% 
 select(Organization, Budget_Size, Current.Final, Scenario, delta, percentChange) %>% 
  kable(digits = 2, format.args=list(big.mark=','), caption="Largest Changes, Scenario6b")

```


```{r}
#all_scenarios %>% mutate(Scenario=Scenario2.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
#  ggplot() + aes(x=percentChange) + geom_histogram()
```


#### Total changes

```{r deltatbl31}
zzz <- all_scenarios[c('Scenario5.Final','Scenario5a.Final','Scenario5b.Final','Scenario6.Final', 'Scenario6a.Final','Scenario6b.Final')] - all_scenarios$Current.Final

zzz <- data.frame(Organization=all_scenarios$Organization, OGP_Budget_Category=all_scenarios$OGP_Budget_Category, zzz)


melt(zzz, c("Organization", "OGP_Budget_Category")) %>% 
  group_by(Scenario=sub(".Final", "", variable), 
           delta=ifelse(sign(value) > 0, '\\+', '\\-')
           ) %>% 
  summarise(n=n(), avg=mean(value), min=min(value), med=median(value), max=max(value)) %>% 
    kable( digits=0, caption="Types / Sizes of Changes")

```

```{r deltatbl231}

melt(zzz, c("Organization", "OGP_Budget_Category")) %>% 
  group_by(Scenario=sub(".Final", "", variable), 
           `Formula Group`=OGP_Budget_Category,
           delta=ifelse(sign(value) > 0, '\\+', '\\-')
           ) %>% 
  summarise(n=n(), avg=mean(value), min=min(value), med=median(value), max=max(value)) %>% 
    kable( digits=0, caption="Types / Sizes of Changes by Formula Group",align='lccrrrrr') %>% 
    row_spec(c(7:13, 20:25, 33:38), background="lightgrey")

```
  



# Integration with website


## Scenario 5 Website collateral
```{r comment=NA}
s1s <- solveForBrackets(scenario5)
cat(to_code(s1s))
```

```{r comment=NA}
cat(to_text(s1s))
```

## Scenario 5a Website collateral
```{r comment=NA}
s1s <- solveForBrackets(scenario5a)
cat(to_code(s1s))
```

```{r comment=NA}
cat(to_text(s1s))
```

## Scenario 5b Website collateral
```{r comment=NA}
s1s <- solveForBrackets(scenario5b)
cat(to_code(s1s))
```

```{r comment=NA}
cat(to_text(s1s))
```

## Scenario 6 Website collateral
```{r comment=NA}
s1s <- solveForBrackets(scenario6)
cat(to_code(s1s))
```

```{r comment=NA}
cat(to_text(s1s))
```

## Scenario 6a Website collateral
```{r comment=NA}
s1s <- solveForBrackets(scenario6a)
cat(to_code(s1s))
```

```{r comment=NA}
cat(to_text(s1s))
```

## Scenario 6b Website collateral
```{r comment=NA}
s1s <- solveForBrackets(scenario6b)
cat(to_code(s1s))
```

```{r comment=NA}
cat(to_text(s1s))
```
