---
title: "OGP 2020 Extra Scenario 6a2"
author: "Neal Fultz <<neal@njnm.co>>"
date: "5/18/2020"
output: html_vignette
vignette: >
  %\VignetteIndexEntry{OGP 2020 Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(ogp)

knitr::opts_chunk$set(echo=FALSE)

if(! "Current.Awards" %in% colnames(allocations)) {
  allocations <- cbind(
    allocations,
    Current = Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year)
  )
}

```






## Scenarios 6a2

```{r scenarios, error=FALSE, results='hide'}

adjustFinalCategoryToConstraint <- function(scen, limit) {
  loss <- function(x) abs(limit - OGP.2013_forward(40000000, 
        scen - c(0,0,0, 0,0,0, x,x,0)))
  x <- optim(0, loss, method="Brent", lower=-.01, upper=scen[8])
  scen - c(0,0,0,0,0,0,x$par,x$par,0)
}




ineq1_cap <- matrix(ncol=10, byrow=TRUE, -c(
1, 7499, 10000, 32500, 50000, 1400000, 8500000, 3e+07, 0, 250000
)/100000)

ineq2_cap <- matrix(ncol=10, byrow=TRUE, -c(
1, 7499, 10000, 32500, 50000, 1400000, 8500000, 3e+07, 0, 200000
)/100000)


ineq1_base <- matrix(ncol=10, byrow=TRUE, c(
1, 0, 0, 0, 0, 0, 0, 0, 0,    6750
))


eq5a <- matrix(ncol=10, byrow=TRUE, c(
0, 1, 0, 0, 0, 0, 0, 0, 0,    0,
0, 0, 1, 0, 0, 0, 0, 0, 0,    0,
0, 0, 0,-1, 1, 0, 0, 0, 0,    0,
0, 0, 0, 0, 0, 0,-1, 1, 0,    0,
0, 0, 0, 0, 0, 0, 0, 0, 1,    0
))





#print(list(s, scenario5b))
#}

# scenario5b <- c(7500, 0, 0, 0.3862, 0.3862, 0.0386, 0.0041, 0.0041, 0)


### Scenario 6 - 100% match for micro up to $7500, recalibrate
eq23 <- matrix(ncol=10, byrow=TRUE, c(
0, 0, -1,  1, 0, 0, 0, 0, 0,   0,
0, 0, -1,  0, 1, 0, 0, 0, 0,   0,
0, 0,  0,  0, 0, 0, 1,-1, 0,   0,  
0, 0,  0,  0, 0, 0, 0, 0,-1,   0
))


ineq23 <- matrix(ncol=10, byrow=TRUE, c(
0, 1, 0,  0, 0, 0, 0, 0, 0,   .99
))


#scenario6 <- c(0, 1, 0.3279, 0.3279, 0.3279, 0.0325, 0.0049, 0.0049, 0)
  
scenario6a <- scenario(allocations,
                      Y_low = .90 * allocations$Current.Grant100,
                      groups_const = c("City", "District_Most_Activity", "Discipline"),
                      mod=list(base=3, extra_eq=eq23, extra_inq=rbind(ineq23,ineq1_cap),
                               s=0.49, method='LU'),
                      niter=82#, verbose=TRUE
                      ); .Last.value
scenario6a[1:2] <- 0:1
#scenario6a <- c(0, 1, 0.3312, 0.3312, 0.3312, 0.0407, 0.004, 0.004, 0)
scenario6a <- adjustFinalCategoryToConstraint(scenario6a, 250000)
scenario6a2 <- adjustFinalCategoryToConstraint(scenario6a, 270000)




```


2. Scenario 6a2 
    - as Scenario 6a, but capped to 270,000

We can evaluate these scenarios at various budget sizes. All scenarios look very similar for the high end, but can be quite different at the small end:

```{r ScenComp}
b <- cumsum(basis(100*1000000))
kable(cbind(Budget=b,
  Scenario6a= OGP.2013_forward(b, scenario6a),
  Scenario6a2= OGP.2013_forward(b, scenario6a2)
), digits = -2, format.args=list(big.mark=',', scientific = FALSE), caption="Max Request Amounts") %>% column_spec(1, border_right = TRUE)

```



```{r error=FALSE, warning=FALSE, fig.height=6, fig.width=8, fig.align="center", out.width="80%"}
zzz <- data.frame(Budget=b,
  Current= OGP.2013_forward(b),
  Scenario6a= OGP.2013_forward(b, scenario6a),
  Scenario6a2= OGP.2013_forward(b, scenario6a2)
)

zzz_long <- melt(zzz, id.vars="Budget", variable.name="Scenario", value.name = "MaxAward")

ggplot(zzz_long) + aes(x=Budget, y=MaxAward, color=Scenario) + geom_line() +
  geom_point(aes(x=Budget_Size), data=allocations, y=0, pch='x', color='black') + 
  scale_y_continuous(name = "Max Request", limits = c(0,300000), labels = scales::dollar) + 
  scale_x_continuous(name = "Budget", limits = c(1,100000000), labels = scales::dollar) +
  ggtitle("Max Request by Budget")

```

```{r error=FALSE, warning=FALSE, fig.height=6, fig.width=8, fig.align="center", out.width="80%"}
# ggplot(zzz_long) + aes(x=Budget, y=MaxAward, color=Scenario) + geom_line(aes(linetype=Scenario != "Current"), size=1.5) + xlim(1, 50000) + ylim(0, 30000) +
#   geom_point(aes(x=Budget_Size), data=allocations, y=0, pch='x', color='black') + guides(linetype="none") + ggtitle("OGP I Scenarios")


ggplot(zzz_long) + aes(x=Budget, y=MaxAward, color=Scenario) + geom_line() +
  geom_point(aes(x=Budget_Size), data=allocations, y=0, pch='x', color='black') + 
  scale_y_continuous(name="Max Request", limits = c(0,30000), labels = scales::dollar) + 
  scale_x_continuous(name="Budget", limits = c(1,50000), labels = scales::dollar) +
  ggtitle("Max Request by Budget (Smaller categories)") + 
  geom_abline(slope=1, intercept=0, linetype='dashed', alpha=.4)

```

# Comparison by final awards (4.5M Total County Budget) by group
```{r all_sec_counterfactual}
  Current  = Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=NULL)
  Scenario6a= Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=scenario6a)
  Scenario6a2= Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=scenario6a2)

  all_scenarios <- cbind( allocations[c('Organization', 'Year', 'OGP_Budget_Category', "District_Most_Activity", "Discipline", "Budget_Size")],
                          Current=Current, 
                          Scenario6a=Scenario6a, 
                          Scenario6a2=Scenario6a2)
  
  write.csv(all_scenarios, "~/extra_scenarios_2.csv")
  
```

### Change in funding rate


```{r}
kable(rbind.data.frame(
  data.frame(Scenario="Current", `Total Scored Request` = sum(Current$Grant100), `Total Final Award`=sum(Current$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`),
data.frame(Scenario="Scenario6a", `Total Scored Request` = sum(Scenario6a$Grant100), `Total Final Award`=sum(Scenario6a$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`),
data.frame(Scenario="Scenario6a2", `Total Scored Request` = sum(Scenario6a2$Grant100), `Total Final Award`=sum(Scenario6a2$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`)
) %>%   mutate(p=round(p,3), `Total Final Award` = NULL, `Total Scored Request` = round(`Total Scored Request`, -2)), 
               format.args=list(big.mark=',')) 
```


#### Largest changes in each scenario

```{r}
all_scenarios %>% mutate(Scenario=Scenario6a.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
  arrange(-Budget_Size) %>% filter(rank(-abs(percentChange))< 20 | rank(-abs(delta))< 20 ) %>% 
 select(Organization, Budget_Size, Current.Final, Scenario, delta, percentChange) %>% 
  kable(digits = 2, format.args=list(big.mark=','), caption="Largest Changes, Scenario6a")

```

```{r}
all_scenarios %>% mutate(Scenario=Scenario6a2.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
  arrange(-Budget_Size) %>% filter(rank(-abs(percentChange))< 20 | rank(-abs(delta))< 20 ) %>% 
 select(Organization, Budget_Size, Current.Final, Scenario, delta, percentChange) %>% 
  kable(digits = 2, format.args=list(big.mark=','), caption="Largest Changes, Scenario6a2")

```


```{r}
#all_scenarios %>% mutate(Scenario=Scenario2.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
#  ggplot() + aes(x=percentChange) + geom_histogram()
```


#### Total changes

We can also aggregate the changes in final reward by whether it increased or not, and by how much:


```{r deltatbl}
zzz <- all_scenarios[c('Scenario6a.Final','Scenario6a2.Final')] - all_scenarios$Current.Final

zzz <- data.frame(Organization=all_scenarios$Organization, OGP_Budget_Category=all_scenarios$OGP_Budget_Category, zzz)


melt(zzz, c("Organization", "OGP_Budget_Category")) %>% 
  group_by(Scenario=sub(".Final", "", variable), 
           delta=ifelse(sign(value) > 0, '\\+', '\\-')
           ) %>% 
  summarise(n=n(), avg=mean(value), min=min(value), med=median(value), max=max(value)) %>% 
    kable( digits=0, caption="Types / Sizes of Changes")

```

```{r deltatbl2}

melt(zzz, c("Organization", "OGP_Budget_Category")) %>% 
  group_by(Scenario=sub(".Final", "", variable), 
           `Formula Group`=OGP_Budget_Category,
           delta=ifelse(sign(value) > 0, '\\+', '\\-')
           ) %>% 
  summarise(n=n(), avg=mean(value), min=min(value), med=median(value), max=max(value)) %>% 
    kable( digits=0, caption="Types / Sizes of Changes by Formula Group",align='lccrrrrr') %>% 
    row_spec(c(8:14), background="lightgrey")

```
  


# Comparison by final awards (3.1M Total County Budget) by group
```{r all_sec_counterfactual31}
  Current  = Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=NULL, TotalBudget = 3100000)
  Scenario6a= Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=scenario6a, TotalBudget = 3100000)
  Scenario6a2= Award(allocations$Budget_Size, OGP.2013_forward, allocations$ScorePercent, Year = allocations$Year, beta=scenario6a2, TotalBudget = 3100000)

  all_scenarios <- cbind( allocations[c('Organization', 'Year', 'OGP_Budget_Category', "District_Most_Activity", "Discipline", "Budget_Size")],
                          Current=Current, 
                          Scenario6a=Scenario6a, 
                          Scenario6a2=Scenario6a2)
  
```

### Change in funding rate


```{r}
kable(rbind.data.frame(
  data.frame(Scenario="Current", `Total Scored Request` = sum(Current$Grant100), `Total Final Award`=sum(Current$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`),
data.frame(Scenario="Scenario6a", `Total Scored Request` = sum(Scenario6a$Grant100), `Total Final Award`=sum(Scenario6a$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`),
data.frame(Scenario="Scenario6a2", `Total Scored Request` = sum(Scenario6a2$Grant100), `Total Final Award`=sum(Scenario6a2$Final), check.names = FALSE) %>% mutate(p=`Total Final Award` / `Total Scored Request`)
) %>%   mutate(p=round(p,3), `Total Final Award` = NULL, `Total Scored Request` = round(`Total Scored Request`, -2)), 
               format.args=list(big.mark=',')) 
```


#### Largest changes in each scenario

Note the Current Column displayed below is the Current OGP Formula with a 3.1M Budget.

```{r}
all_scenarios %>% mutate(Scenario=Scenario6a.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
  arrange(-Budget_Size) %>% filter(rank(-abs(percentChange))< 20 | rank(-abs(delta))< 20 ) %>% 
 select(Organization, Budget_Size, Current.Final, Scenario, delta, percentChange) %>% 
  kable(digits = 2, format.args=list(big.mark=','), caption="Largest Changes, Scenario6a")

```

```{r}
all_scenarios %>% mutate(Scenario=Scenario6a2.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
  arrange(-Budget_Size) %>% filter(rank(-abs(percentChange))< 20 | rank(-abs(delta))< 20 ) %>% 
 select(Organization, Budget_Size, Current.Final, Scenario, delta, percentChange) %>% 
  kable(digits = 2, format.args=list(big.mark=','), caption="Largest Changes, Scenario6a2")

```


```{r}
#all_scenarios %>% mutate(Scenario=Scenario2.Final, delta=Scenario - Current.Final, percentChange= 100*(Scenario /Current.Final - 1)) %>% 
#  ggplot() + aes(x=percentChange) + geom_histogram()
```


#### Total changes

```{r deltatbl31}
zzz <- all_scenarios[c('Scenario6a.Final','Scenario6a2.Final')] - all_scenarios$Current.Final

zzz <- data.frame(Organization=all_scenarios$Organization, OGP_Budget_Category=all_scenarios$OGP_Budget_Category, zzz)


melt(zzz, c("Organization", "OGP_Budget_Category")) %>% 
  group_by(Scenario=sub(".Final", "", variable), 
           delta=ifelse(sign(value) > 0, '\\+', '\\-')
           ) %>% 
  summarise(n=n(), avg=mean(value), min=min(value), med=median(value), max=max(value)) %>% 
    kable( digits=0, caption="Types / Sizes of Changes")

```

```{r deltatbl231}

melt(zzz, c("Organization", "OGP_Budget_Category")) %>% 
  group_by(Scenario=sub(".Final", "", variable), 
           `Formula Group`=OGP_Budget_Category,
           delta=ifelse(sign(value) > 0, '\\+', '\\-')
           ) %>% 
  summarise(n=n(), avg=mean(value), min=min(value), med=median(value), max=max(value)) %>% 
    kable( digits=0, caption="Types / Sizes of Changes by Formula Group",align='lccrrrrr') %>% 
    row_spec(c(8:14), background="lightgrey")

```
  



# Integration with website

## Scenario 6a Website collateral
```{r comment=NA}
s1s <- solveForBrackets(scenario6a)
cat(to_code(s1s))
```

```{r comment=NA}
cat(to_text(s1s))
```

## Scenario 6a2 Website collateral
```{r comment=NA}
s1s <- solveForBrackets(scenario6a2)
cat(to_code(s1s))
```

```{r comment=NA}
cat(to_text(s1s))
```
